---
title: "get maps"
author: "Alex(ks)'s Bushby and Jovic"
date: "March 20, 2018"
output: html_document
---

##packages

```{r packages}
library(ggplot2)
library(tidyverse)
library(forcats)
library(labeling)
library(ggmap)
```


#ggmap

* ggmap easily gets raster map tiles from online sources (raster data: simplest form, a raster consists of a matrix of cells (or pixels) organized into rows and columns (or a grid) where each cell can be assigned a value (such as temperature, precipitation, etc).
* These maps can be digital aerial photographs, imagery from satellites, digital pictures, or even scanned maps.
* Plots are done then using the ggplot2 framework (where you identify your geom layers).
* ggmap also has functions for geolocation and routing.


## Getting Maps

Before we do anything else, we need our map layer upon which everything else will be plotted.

ggmap has a variety of functions (which have varying parameters) which can pull up maps from popular sources: Google Maps, Stamen Maps, OpenStreetMap

## Retrieving StamenMap

Suppose you want a plain map of North America:

```{r stamenmap}
mapzone = c(left = -135, bottom = 25.75, right = -62, top = 75) #you specify the box
map = get_stamenmap(mapzone, zoom = 5, maptype = "toner-lite") 
ggmap(map) #prints the map for viewing
```

Note: large ggmaps can grow massively in size (150MB) since they are pieced together

## Retrieving GoogleMap

When retreiving with googlemap, here we only need to center it.

```{r googlemap}

map1 = get_googlemap(center=c(lon=-79.91, lat=43.26),zoom=13, size=c(640,640),scale=2,maptype='terrain') #'satellite','roadmap','hybrid'  are the other maptype offered

ggmap(map1) #hamilton for example, lon,lat very sensitive, so need to convert hours/minutes into a decimal
#zoom must be whole number, so a bit finnicky

#mcmaster satellite map
mcmastermap = get_googlemap(center=c(lon=-79.9192, lat=43.2609), zoom=15, maptype = 'satellite'); ggmap(mcmastermap)

#you can get as close as the database permits
fireball = get_googlemap(center=c(lon=-79.92, lat=43.2609), zoom=19, maptype = 'satellite'); ggmap(fireball)
```
