---
title: "2moremaps"
author: "Alex(ks)'s Bushby and Jovic"
date: "March 20, 2018"
output: html_document
---


##packages

```{r packages}
library(ggplot2)
library(tidyverse)
library(forcats)
library(labeling)
library(ggmap)
```


## Over Querey Limit exceptions

One issue that came up was getting the exception: "Over Query Limit"... even if it's your first time running it.

StackOverFlow suggested a solution:

ggmap v2.7 allows the user to specify a Google Maps API key through the  register_google() function. v2.7 is not on CRAN yet, so you have to use devtools::install_github("dkahle/ggmap") to install it. After updating to that version and setting my API Key register_google(key = "my_api_key"), the same batch worked in one run as expected.

The Google Maps API Key is easy to get: https://developers.google.com/maps/documentation/geocoding/get-api-key  (credit to user: sbha)

## Credentials

ggmap will set credentials to the following
ggmap_credentials()
#  Google - 
#     key :  
#     account_type : standard 
#     day_limit : 2500 
#     second_limit : 50 
#     client :  
#     signature :

If you have a Google API key, you could get more, but you would need to register.
register_google(key = "[your key here]", account_type = "premium", day_limit = 10000)
This will then be checked. Clearly this is an inconvenience, but shouldn't be a problem for regular use.

## More quick methods with googlemaps

If you'd just like to quickly zoom in on downtown, the following will quickly do the trick.
```{r moregoog}
#can also just type city (for american cities:
#nested (traditional) look
ggmap(get_googlemap("london england", zoom = 12))

#with piping
get_googlemap("newyork newyork", zoom = 12) %>% ggmap() #gives me las vegas?
get_googlemap('toronto canada', zoom=12,maptype='satellite') %>% ggmap() #certain zooms won't 12 does, 13 doesnt, 14 doesnt, 15 does
get_googlemap('toronto canada', zoom=15, maptype='hybrid') %>% ggmap() #hybrid gives names of landmarks/streets with the satellite
```
