---
title: "3utility"
author: "Alex(ks)'s Bushby and Jovic"
date: "March 20, 2018"
output: html_document
---
##packages

```{r packages}
library(ggplot2)
library(tidyverse)
library(forcats)
library(labeling)
library(ggmap)
```

# ggmap Utility Functions

## geocode()

Suppose you want to collect location data (if you're assembling a dataset, or just in general).
Geocode takes an address and returns coordinates.
```{r geolookup}
geocode("mcmaster university", output="more")
```
output="more" can be specified if you want more extensive information other than just the lat/lon

## revgeocode()

You can also use this package as an atlas of sorts:
```{r reverse}
coordinates = geocode("mcmaster university")%>%as.numeric()

revgeocode(coordinates)
```

## Distances

We can also measure map distances using mapdist(from='x',to='y').

Lesser known, or cities that don't have a unique name require the country (in many cases).

```{r distance}
mapdist(from="toronto",to="hamilton") #observe, won't work

x=c('toronto','boston'); y=c('hamilton, canada','cleveland')
mapdist(x,y)
```

## Route planning (the hard way)

We can use route() or trek(). trek() tends to hug the roads more closely.

```{r route}
routeplan= route(
  from='mcmaster university',
  to='guelph university',
  alternatives= TRUE
)
qmap('puslinch, canada', zoom = 10, maptype = 'roadmap',
     base_layer=ggplot(aes(x=startLon,y=startLat),data=routeplan))+
  geom_leg(
    aes(x=startLon,y=startLat,xend=endLon,yend=endLat,
           colour=route),
alpha=.7, size=2, data=routeplan) +
  labs(x="longitude",y="latitude",colour="Route")+
  facet_wrap(~route,ncol=2)+
  theme(legend.position='bottom')
```

Here we used ggplot for the first time, by adding layers to qmap. We will explore it in more detail in the next section.
