---
title: "4crime example"
author: "Alex(ks)'s Bushby and Jovic"
date: "March 20, 2018"
output: html_document
---
##packages

```{r packages}
library(ggplot2)
library(tidyverse)
library(forcats)
library(labeling)
library(ggmap)
```


# Visualizing spatial data over a map (Putting it all together)

Pulling up maps in R is a fun distraction, but let's put it together with some data.

## qmplot sidenote

ggmap likes to build its plots in a function called qmplot which is the ggplot equivalent of qplot. Is that a bad thing? It appears that's the way things are done with this package.

qplot is known as the "quick and dirty" way to quickly produce graphs, but it runs into issues when you try to do more complicated things.

example below of qplot vs ggplot (user kohske on stackoverflow)

qplot(x,y, geom="line") # I will use this
ggplot(data.frame(x,y), aes(x,y)) + geom_line() # verbose

d <- data.frame(x, y)

qplot(x, y, data=d, geom="line") 
ggplot(d, aes(x,y)) + geom_line() # I will use this

qmap is a "wrapper" for ggmap and get_map: qmap(location = "X", ...)

## Example with the crime dataset

The ggmap package has a 'crime' dataset built in; the dataset contains the crimes for Houston from Januaray to August (2010), geocoded with Google Maps.

```{r crimedata}
head(crime)
#View(crime) #for full
```
We can see the data contains a lot of useful info: dates, types of crimes, locations by type of place, locations by street, locations by longitude/latitude.

First, let's get an overview of the crimes on the map.
Using qmplot, we put in longitude and latitude for the x, and y parameters, and specify the data as the crime dataset. This plots the crimes indiscriminantly.
```{r crime1}
# to see all crimes
qmplot(lon, lat, data = crime, maptype = "toner-lite", colour = I("red"),size = I(0.9),alpha=.3) +
  theme(legend.position="none") #to remove the .3 from the legend (pointless)
```

## Working with the data

Suppose we are interested in violent crimes in the downtown Houston area.

A slick method to isolate this could involve dplyr to create a 'new pipe' (similar to defining a new class method)

```{r helper}
`%notin%` = function(setA, setB) !(setA %in% setB) #credit to kahle, checks that setA is not in setB

# reduce crime to violent crimes in downtown houston
violent_crimes = crime %>% filter(
    offense %notin% c("auto theft", "theft", "burglary"),
    -95.39681 <= lon & lon <= -95.34188,
    29.73631 <= lat & lat <=  29.78400
  ) %>% 
  mutate(
    offense = fct_drop(offense), #drops unused levels, does not drop NA levels that have values, will prevent errors
    offense = fct_relevel(offense,  #reorders factor levels by hand
                          c("robbery", "aggravated assault", "rape", "murder")
    )
  )

## Here is a traditional way to code it without pipes:
#
# qmplot(lon, lat, data = crime)
# 
# 
# # only violent crimes
# violent_crimes = subset(crime,
#   offense != "auto theft" &
#   offense != "theft" &
#   offense != "burglary"
# )
# 
# # rank violent crimes
# violent_crimes$offense = factor(
#   violent_crimes$offense,
#   levels = c("robbery", "aggravated assault", "rape", "murder")
# )
# 
# # restrict to downtown
# violent_crimes = subset(violent_crimes,
#   -95.39681 <= lon & lon <= -95.34188 &
#    29.73631 <= lat & lat <=  29.78400
# )
# 
# theme_set(theme_bw())
# 
# qmplot(lon, lat, data = violent_crimes, colour = offense,
#   size = I(3.5), alpha = I(.6), legend = "topleft")
# 
# qmplot(lon, lat, data = violent_crimes, geom = c("point","density2d"))
# qmplot(lon, lat, data = violent_crimes) + facet_wrap(~ offense)
# qmplot(lon, lat, data = violent_crimes, extent = "panel") + facet_wrap(~ offense)
# qmplot(lon, lat, data = violent_crimes, extent = "panel", colour = offense, darken = .4) +
#   facet_wrap(~ month)

qmplot(lon, lat, data = violent_crimes, maptype = "toner-lite", colour = I("red"),size = I(0.9),alpha=.3) +
  theme(legend.position="none")
```

## Different visualizations

In the previous example, the geom parameter is set by default to "point". Let's see it with the countour and bin geometries:
```{r}
qmplot(x=lon, y=lat, data = violent_crimes, maptype = "toner-2011", geom = "density2d", colour = I("red"),size=I(1))

qmplot(x=lon, y=lat, data = violent_crimes, maptype = "toner-2011", geom = "bin2d")
```
Neither of these graphs are particularly great, so let's see if we can improve on them.
The countour is a bit hard to tell what areas have higher crime, and the bins are differentiated by hue, but it clashes a bit with the map itself. Trying to colour the bin puts a border around it, which only makes the hue harder to distinguish (optical illusion).

## Heatmap

To make the countour map more useful, we can assign a gradient. Let's look at the robberies:

```{r robberies}

robberies = violent_crimes %>% filter(offense=='robbery')

qmplot(lon, lat, data = violent_crimes, geom = "blank", 
       zoom = 15, maptype = "toner-background", legend = "bottomright"
) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = .35, colour = NA) +
  scale_fill_gradient2("Robbery\nHeatmap", low = "white", mid = "yellow", high = "red", midpoint = 650)

```

## Faceting different categories, and not using qmplot

We can also look at different things, depending on what we're interested in:

```{r facets}

#faceting by crime
(
qmplot(lon, lat, data = violent_crimes, maptype = "toner-background", colour = offense) + 
  facet_wrap(~ offense)
)

#faceting by day

#First specify your map
houston = get_googlemap('houston texas', zoom=15, maptype='roadmap')
ggmap(houston)

#then add the layers like we would in ggplot
HoustonMap = ggmap(houston, base_layer = ggplot(aes(x = lon, y = lat),
                                                 data = violent_crimes))
HoustonMap +
  stat_density2d(aes(x = lon, y = lat, fill = ..level.., alpha = ..level..),
                 bins = 5, geom = "polygon",
                 data = violent_crimes) +
  scale_fill_gradient(low = "black", high = "red") +
  facet_wrap(~ day) +
  theme(legend.position="none")
```

The map we supplied here (roadmap with google) is probably a bit too colourful, but this is just to illustrate the capabilities.

The following was VERY experimental, and I couldn't get it to work out properly... perhaps too many locations, but the idea is that we can include more information by adding an additional piece to the legend --> We let the offense be coded by colour, and the goal was to let the type of location be denoted by a particular symbol.

```{r try}
theme_set(theme_bw())
houmap = qmap("houston",zoom=13, colour='bw', legend ='topleft')

(houmap+
    geom_point(data=violent_crimes, aes(x=lon,y=lat,colour=offense,shape=location))
)

(houmap+
    geom_point(data=violent_crimes, aes(x=lon,y=lat,colour=offense,shape=location))+
    theme(legend.position="none")
)
```

We can also put the data into bins if we are interested in an overview (if we are more interested in zones).

```{r bin}
(
houmap+
    stat_bin2d(
      data=violent_crimes,
      aes(x=lon,y=lat,colour=offense,fill=offense),
      size=0.6, bins= 90, alpha=.4
    )
)
#compare to closer, we can use less bins
houmap = qmap("houston",zoom=14, colour='bw', legend ='topleft')

(houmap+
    geom_point(data=violent_crimes, aes(x=lon,y=lat,colour=offense))
)

(
  houmap+
    stat_bin2d(
      data=violent_crimes,
      aes(x=lon,y=lat,colour=offense,fill=offense),
      size=0.6, bins= 30, alpha=.4
    )
)

```
Depending on the level of detail we want, we can manually adjust the zoom, and the bin number.

Note, it is important that as you zoome in, you'll want to reduce the number of bins, or you may as well just use the dot-plot since each crime will fit uniquely in its own bin.
