---
title: "Spatial Data"
author: "The Alex(k)'s"
date: "3/14/2018"
output: html_document
---

```{r, echo=FALSE}
library(shiny)
```


Maps allow us to locate objects geographically and the added features (such as shape, size and colour) can provide insight to characteristics of spatial data^1^



##Three Strategies ^1^
1. Display
2. Analyze
3. Explore

Consider audience, interactivity and nature of data.


##Seven Techniques^2^

1. Chloropleth Map: different colours/shadings for different regions
2. Heat Map: similar to a chloropleth map, but is not defined by regions
3. Hexagonal Binning: use regular hexagons to create bins in your map
4. Dot Map: scatterplots on a map
5. Cluster Map: scatterplots on a map, where each point represents the number of points grouped together
6. Bubble Map: colour and size can represent two different variables
7. Cartogram Map: regions of a map represent a variable



##Things to Consider
(and we will return to these items throughout our prsentation)

* given a large dataset, how much of the data should we display?^3^
* what colours should be used for the maps and for the information held in the maps?
* should we bin information or should we have a continuous colour scale?
* when binning, how do we choose an appropriate size?
* when to use a contour plot versus a fill plot?




##Highly Recommend this Talk!

[Google Maps + HTML5 + Spatial Data Visualization: A Love Story](https://www.youtube.com/watch?v=aZJnI6hxr-c)

#Packages Required

##Leaflet
```{r, message=FALSE}
library(leaflet)
library(knitr)
library(rgdal) #issues on windows, created for OSX
```

##ggmap
```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
library(forcats)
library(labeling)
library(ggmap)
```

#ggplot
See ggplot rmd

#Leaflet

[Leaflet for R](https://rstudio.github.io/leaflet/) is where I learned how to use Leaflet. It is a great resource, so refer to it if you need more clarification.

`addTiles()` gives the default base map in leaflet^4^

```{r}
m <- leaflet() %>% 
  addTiles()
m
```

##Base Maps

`addProviderTiles()` gives you a tile layer from a known map provider^4^

```{r}
m %>% addProviderTiles(providers$Stamen.Toner) 
```

```{r}
m %>% addProviderTiles(providers$CartoDB.Positron)
```

```{r}
m %>% addProviderTiles(providers$Esri.NatGeoWorldMap)
```


```{r, message=FALSE, echo=FALSE}
#install.packages("kableExtra")
#devtools::install_github("haozhu233/kableExtra")

library(kableExtra)
library(knitr)

text_tbl <- data.frame(
  BaseMap = c("Default", "Stamen.Tower", "CartoDB.Positron", "Esri.NatGeoWorldMap"),
  Pros = c("Zoomed out - colours are fairly simple", "None?", "Great for added colour as the background colour is mutual", "Shows different landscapes of Earth, i.e. deepness of water & dry/wet regions"),
  Cons = c("Zoom in - gives extra colour", "Any colour added to plot will likely be hard to see", "None?", "Specific uses only")
)

kable(text_tbl, "html") %>%
  kable_styling(full_width = T) 
```


###Example: Base Map versus CartoDB.Positron Map (Weather Systems)

`addWMSTiles` is for adding Web Map Service tiles^4^

```{r}
leaflet() %>% addTiles() %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad"
  )
```

```{r}
leaflet() %>% addProviderTiles(providers$CartoDB.Positron) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad"
  )
```

##Circle Markers

```{r}
data(quakes)

leaflet(data = quakes) %>%  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~long, ~lat, radius = ~sqrt(exp(mag))/2, label = ~as.character(mag), color = "black", fillColor = "red", fillOpacity = 0.1, stroke = FALSE, opacity = 0)

leaflet(data = quakes) %>%  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~long, ~lat, radius = ~sqrt(exp(mag))/2, label = ~as.character(mag), color = "black", fillColor = "red", fillOpacity = 0.6, stroke = FALSE, opacity = 0)
```

**Comments about the plot above:**

* when you zoom in, there is more detail shown
* option to make the fillOpacity small so the patterns can be more distinguishable
* problem: can't distinguish between different magnitudes (which are represented in the size)
* question: how should I scale the magnitude (as magnitude is in a log scale) and what colour should I represent to show earthquakes?
* next step: find a way to represent plate boundaries to create another layer over this map.

##Read OGR (OpenGIS Simple Features Reference Implementation)

```{r}
library(rgdal)
states <- readOGR("shapefiles/cb_2016_us_state_20m.shp",
 layer = "cb_2016_us_state_20m", GDAL1_integer64_policy = TRUE)

leaflet(states) %>%
    addPolygons(color = "#444444", weight = 2, smoothFactor = 0.5,
                opacity = 0.1, fillOpacity = 0.8,
                fillColor = ~colorQuantile("YlOrRd", ALAND)(ALAND),
                highlightOptions = highlightOptions(color = "white", weight = 2,
                                                    bringToFront = TRUE))

leaflet(states) %>% addProviderTiles(providers$CartoDB.Positron) %>%
    addPolygons(color = "#444444", weight = 2, smoothFactor = 0.5,
                opacity = 0.1, fillOpacity = 0.8,
                fillColor = ~colorQuantile("YlOrRd", ALAND)(ALAND),
                highlightOptions = highlightOptions(color = "white", weight = 2,
                                                    bringToFront = TRUE))
```

**Comment about the plot above:**

* question: when should one include `\addProviderTiles()` or `\addTiles()`? 

##Chloropleth Maps

```{r}
tmp <- tempdir()
url <- "http://personal.tcu.edu/kylewalker/data/mexico.zip"
file <- basename(url)
download.file(url, file)
unzip(file, exdir = tmp)
mexico <- readOGR(dsn = tmp, layer = "mexico", encoding = "UTF-8")
pal <- colorBin("Reds", mexico$gdp08, 6, pretty = FALSE) ##6 is the number of bins
state_popup <- paste0("<strong>Estado: </strong>", 
                      mexico$name, 
                      "<br><strong>PIB per c?pita, miles de pesos, 2008: </strong>", 
                      mexico$gdp08)
leaflet(data = mexico) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(gdp08), 
              fillOpacity = 0.8, 
              color = "grey", 
              weight = 2, 
              dashArray = "3", 
              highlight = highlightOptions(weight = 5, 
                                            
                                           bringToFront = TRUE),
              label = ~as.character(mexico$state))
```
**Comments on the plot above**

* The bin size needs to be organized better

* It would be good to add the value of the density to the label

* Colour is important here - started with blue and it wasn't very helpful as blue can be mixed with the grey in the background

* Legend would be very useful and positioning of legend is important

```{r}
tmp <- tempdir()
url <- "http://personal.tcu.edu/kylewalker/data/mexico.zip"
file <- basename(url)
download.file(url, file)
unzip(file, exdir = tmp)
mexico <- readOGR(dsn = tmp, layer = "mexico", encoding = "UTF-8")
bin<-c(0,40,60, 80, 100, Inf)
pal <- colorBin("Reds", mexico$gdp08, bin, pretty = FALSE) ##6 is the number of bins
state_popup <- paste0("<strong>Estado: </strong>", 
                      mexico$name, 
                      "<br><strong>PIB per c?pita, miles de pesos, 2008: </strong>", 
                      mexico$gdp08)
leaflet(data = mexico) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(gdp08), 
              fillOpacity = 0.8, 
              color = "grey", 
              weight = 2, 
              dashArray = "3", 
              highlight = highlightOptions(weight = 5, 
                                            
                                           bringToFront = TRUE),
              label = ~as.character(mexico$state)) %>% addLegend(pal = pal, values = ~gdp08, opacity = 0.7, title = "GDP",
  position = "topright")
```

**Next steps: Continuous colour**

```{r}
tmp <- tempdir()
url <- "http://personal.tcu.edu/kylewalker/data/mexico.zip"
file <- basename(url)
download.file(url, file)
unzip(file, exdir = tmp)
mexico <- readOGR(dsn = tmp, layer = "mexico", encoding = "UTF-8")
pal <- colorNumeric("Reds", mexico$gdp08) ##6 is the number of bins
state_popup <- paste0("<strong>Estado: </strong>", 
                      mexico$name, 
                      "<br><strong>PIB per c?pita, miles de pesos, 2008: </strong>", 
                      mexico$gdp08)
leaflet(data = mexico) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = ~pal(gdp08), 
              fillOpacity = 0.8, 
              color = "grey", 
              weight = 2, 
              dashArray = "3", 
              highlight = highlightOptions(weight = 5, 
                                            
                                           bringToFront = TRUE),
              label = ~as.character(mexico$state)) %>% addLegend(pal = pal, values = ~gdp08, opacity = 0.7, title = "GDP",
  position = "topright")
```
**We have the same problem as before with the first discrete colour plot**

##References

[1] Kraak, M.J., "Visualising spatial distributions." In P.A. Longley, &. [et al](Eds.), *Geographical information systems: priciples, techniques, management and applications: abridged*. (2005). Retrieved from https://pdfs.semanticscholar.org/36e4/8a926046f02f42fb7e18aa50d10fbf8d1f11.pdf

[2] Ayalasomayajula, V., "7 Techniques to Visualize Geospatial Data." *Social Cops*. (2016) . Retrieved from https://blog.socialcops.com/academy/resources/7-techniques-to-visualize-geospatial-data/

[3] Zheng, Y., Ou, Y., Lex, A., & Phillips, J.M., "Visualization of Big Spatial Data using Coresets for Kernel Density Estimates." *Cornell University*. (2017). Retrieved from https://arxiv.org/abs/1709.04453

[4] Leaflet for R., "Leaflet." *R Studio*. (2016). Retrieved from https://rstudio.github.io/leaflet/

[5] 
```{r cite, echo=FALSE}
citation('ggmap')
```

##Addendum: Some more advanced topics on Spatial Statistics for future reading.

Here is a package that I didn't have time to full get into: https://cran.r-project.org/web/packages/spdep/spdep.pdf
It provides tools for analyzing spatial dependence: weighting schemes, and models.

I came across an analytical course on applied spatial statistics, but this is a whole course on its own, so I'll just post the link should anyone be interested in the future:
https://scholar.harvard.edu/zhukov/classes/applied-spatial-statistics-r