---
title: "Lecture Jan23"
author: "Aleksandar Jovic"
date: "January 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###

FIRST HALF {Individual variables, Bivariate data}

Banking refers to pool, where 90 degrees is easier. Recall Rauser, important data's angles should be stretched so most of it is at 90 degrees (easier to visualize).

Showed smokers having stroner lungs is misleading since adults have stronger lungs than children, so would need to differentiate that.

In a scatterplot where points stack, can remedy this with:

* jittering: (scattering points a bit, while may be misleading, if it's understood, helps spread stacked points).
* transparency: using alpha=0.4 (or whatever)
* area: but be careful with area, since ggplot will show scale 1 to 2, but 1 to 2 may not be exactly twice as large (i.e. disconnected from PHYSICAL representation of real life)

ggplot uses quadratic LOESS smoother:---->   gg_smooth()

* Once we show this loess trend after differentiating smokers/nonsmokers, we can see smokers aren't actually stronger lungs than non smokers (see plot)
* faceting for sex gives even more info (i.e. comparison btwn sex obviously)

LOESS is local, but not robust.
LOESS function has a choice of family (Gaussian vs sensitivity). Sensitivity claims robust alternative. In geom.smooth() there is an argument to pass how the function works... look into later

With newer technology, we could get a wireframe, which helps us view data in 3D.

When looking at contour plot:

* Darker values being lower, and lighter values higher is awkward
* Continuous scale, showing discrete regions is odd
* Adding dark lines around teach zone makes things very clear (since hue can be hard to distinguish)
* Anchoring y at 0 would help since it's physical quantity would be nice here.

Hexbin fast algorithm (good for large sets).

Colour principles in notes.

SECOND HALF {Multiple dimensions, Multiple factors}

```{r library preamble, echo=TRUE}
library(pgmm)
library(skimr)
library(ggplot2); theme_set(theme_bw())
library(GGally)
library(corrplot)
library(agridat)
library(dplyr)
data(olive)
```

* Joining region data.
* pairs plot
* ggpairs
* convert area, region back to factors
* keep factors in reading order?

Big data: number of observations

* javascript? D3 js?
* summarize first: contour vs hexbin, loess vs gam

High-dimensional: Multiway, hypervariate

```{r example, echo=TRUE}
library(sos)
findFn('barley')
?RSiteSearch
help.search('barley')
fisher.barley

gg0 = (ggplot(fisher.barley)+
         geom_boxplot(aes(x=gen,y=yield,fill=factor(year)))
  
)
print(gg0) #notice, bars will DODGE automatically

#See what happens when we add "notch=TRUE"
gg0a = (ggplot(fisher.barley)+
         geom_boxplot(aes(x=gen,y=yield), #factor year taken out
                      notch=TRUE) #not the best graph for notching
  
)
print(gg0a)

gg0b = (ggplot(fisher.barley,aes(x=gen,y=yield,colour=env))+
         #geom_boxplot(aes(x=gen,y=yield,fill=factor(year)))
          geom_point()+geom_line()+
          facet_wrap(~year)
)
print(gg0b)

ff= (fisher.barley
     %>% mutate(gen=reorder(gen,yield),
                env=reorder(env,yield),
                year=factor(year))
)
gg0c = (ggplot(ff,
               aes(x=env,y=yield,colour=year, group=year))+
         #geom_boxplot(aes(x=gen,y=yield,fill=factor(year)))
          geom_point()+
          geom_line()+
          facet_wrap(~gen,nrow=1)
)

print(gg0c)
```

The function windows() is the microsoft equivalent to quartz() ... this puts plots out into a new window.

Example:

```{r windows fn, echo=TRUE}
windows(); #quartz() #for mac
print(gg0c)
```

Moving onto next example:

```{r olive example} 
#remember, echo=TRUE is the default.. i.e. shows the code
data(olive)
#bolker not a fan of throwing out region info by turning them into numbers, when they could have just been left as factors... side note
# he will push the csv to the repo so we can use the olive data set in a more useful way (keeping information so its not lost and relegated to the info file)
olive_regions= read_csv("Data/olive_regions.csv")  ### NEED TO FIX THIS WORKING DIRECTORY AFTER GETTING THE DATASET FROM BOLKER'S REPO

olive2=(full_join(olive,olive_regions,
          by=c("Region"="region_num",
                "Area"="area_num"))
        %>% select(-c(Region,Area))
) #could make everything pipey, but fine for now

colvec=c("red","blue","orange")
pairs(olive2[,1:8],gap=0,col=colvec[as.numeric(olive2$region)]) 
#gap=0 removes the empty space which Tufte argues is wasted and detracts from data

#now try GGpairs version
ggpairs(olive2,
        columns=1:8,
        mapping=aes(colour=region))

corrplot(cor(olive2[,1:8]),method="ellipse")
corrplot.mixed(cor(olive2[,1:8]))

theme(panel.spacing=grid::unit(0,"lines"))+
  scale_colour_brewer(palette="Dark2")
#recall, our theme is set to black and white, typically will get colour as factors here
```