---
title: "Lecture Jan23"
author: "Aleksandar Jovic"
date: "January 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###

FIRST HALF {Individual variables, Bivariate data}

Banking refers to pool, where 90 degrees is easier. Recall Rauser, important data's angles should be stretched so most of it is at 90 degrees (easier to visualize).

Showed smokers having stroner lungs is misleading since adults have stronger lungs than children, so would need to differentiate that.

In a scatterplot where points stack, can remedy this with:

* jittering: (scattering points a bit, while may be misleading, if it's understood, helps spread stacked points).
* transparency: using alpha=0.4 (or whatever)
* area: but be careful with area, since ggplot will show scale 1 to 2, but 1 to 2 may not be exactly twice as large (i.e. disconnected from PHYSICAL representation of real life)

ggplot uses quadratic LOESS smoother:---->   gg_smooth()

* Once we show this loess trend after differentiating smokers/nonsmokers, we can see smokers aren't actually stronger lungs than non smokers (see plot)
* faceting for sex gives even more info (i.e. comparison btwn sex obviously)

LOESS is local, but not robust.
LOESS function has a choice of family (Gaussian vs sensitivity). Sensitivity claims robust alternative. In geom.smooth() there is an argument to pass how the function works... look into later

With newer technology, we could get a wireframe, which helps us view data in 3D.

When looking at contour plot:

* Darker values being lower, and lighter values higher is awkward
* Continuous scale, showing discrete regions is odd
* Adding dark lines around teach zone makes things very clear (since hue can be hard to distinguish)
* Anchoring y at 0 would help since it's physical quantity would be nice here.

Hexbin fast algorithm (good for large sets).

Colour principles in notes.

SECOND HALF {Multiple dimensions, Multiple factors}



```{r cars, echo= TRUE}
1+2
```
