---
title: "Jan29 Lecture"
author: "Aleksandar Jovic"
date: "January 29, 2018"
output: html_document
---

#if notes posted on time just add comments to .Rmd posted
<https://github.com/bbolker/stat744/blob/master/lectures/diagnostics.rmd>

```{r setup, include=TRUE}
library(tidyverse)
theme_set(theme_bw())
library(car)
library(broom)
library(magrittr)
library(lme4); library(MCMCglmm); library(glmmTMB) ## for models
library(coda) ## for trace plots
library(lattice) ## built-in
library(cowplot)
library(nullabor)
```
Diagnostics: principles

* Don't make decisions on p-values: data collected in the real world is almost certainly not normal, so checking p-value checks if it's normal or not.
* If you do a diagnostic plot, you know nothing is ever EXACTLY linear or normal or equal variances etc... but are the deviations large enough that they'll change the conclusions in a menaingful way (all of this sounds 'waffly')

Diagnostics after fitting model

* qq plot didnt look great, but it was plotting marginals that were normal (see code in Rmd file from note)
* code example bleow, missing dd, see rmd

```{r quantile plots}
m0=lm(y~x,dd)
augment(m0) %>% select(x,y,.resid) %>%
  gather(type,value) %>%
  mutate(type=factor(type,levels=c('x','y','.resid'))) -> a0

ggplot(a0,aes(sample=value))+
  stat_qq()+
  facet_wrap(~type)+
  stat_qq_line(colour='red')
```

Model diagnosis

* mean, variance, distribution (in order of decreasing importance), that is bias, heteroscedasticity, non-normality


Bias (using diamonds dataset from ggplot (?))

```{r bias}
m1=lm(price~carat,diamonds)
a1=augment(m1,data=diamonds) ## include original data
ggplot(a1,aes(.fitted,.resid))+
  geom_point(alpha=0.1)+
  geom_smooth()
```

bias 2 : faceting/colouring

* can change colour of fitted and residual in our model to see where the patterns are coming from

Solutions to bias

* fitting non-linear pattern (recall Box-Cox transform? probably better to use acepack::avas)
* however this gives one or two knobs to adjust, while splines can perhaps give a more complex look
* could also use link functions when needing to develop good models

Heteroscedasticity

* log transformation should be first thing to try and fix (if error variance proportional to mean)
* confidence intervals will fan as you go away from center, doesnt have anything to do with problem in the model, can standarize these residuals in a way to get rid of that pattern
* common for LOESS to get dragged down by extreme point

Quantile plots
* bimodality creates steps
* qqplot draws line through quartiles, not a regression to not get skewed at tails

